---
title: "The Science behind Astrology"
author: "Tesfatsion Shiferaw"
date: "4/24/2020"
output: html_document
---

```{r loadthings, echo=FALSE, message=FALSE}
library(mosaic)
library(ggplot2)
library(gplots)
library(dplyr)
library(ggpubr)
library(RColorBrewer)
#library(fivethirtyeight)
#library(plyr)
```

```{r import}
#DO NOT FORGET FEB.29
patients_data <- read.csv("C:/Users/tesfa/Desktop/Creationof_Stats/Final project/patients_data.csv")
prison_data <- read.csv("C:/Users/tesfa/Desktop/Creationof_Stats/Final project/prisoners_data.csv")
athletics_data <- read.csv("C:/Users/tesfa/Desktop/Creationof_Stats/Final project/athletes_data.csv")
```

## Data clean up

For the first and second question on prison data I did the following data clean up:
*I created a new column of BMI from the height and column of the rows using the BMI formula mass/height^2 (lb/inches^2)
*I changed the date format (birhtdate column) and extract the year out of it so that I can use people of the same age group.
*for the second question I filtered only those rows with offender statues of parole and absconder.
*I removed all the weight less than 97.5lbs because 1. IQR test labled it as an outlier and 2. I assume that people who go to prison are maily adults; hence I believe their weight should be at least greater than 97.5 pounds (which is too small for an adult). I also removed the height less than 53.5 inches for the same reason. ## try removing some more stuff here (BMI>50)

Data clean up for athletics dataset:
*I removed those whose birthdates are after 2003. The youngest athlete to take part in Rio olympics is Gaurika Singh (13 years old) who was born in November 2002. Hence, the rows with birthdates after 2003 are invalid.

```{r data check, filter and clean up}
prison_data['BMI'] <- 703*(prison_data$weight/(prison_data$height*prison_data$height)) #formula of BMI for lb/inches^2
#mean(prison_data$BMI, na.rm=TRUE)
pr_bd <- as.Date(prison_data$date_of_birth, format="%Y-%m-%d") #convert the date format and extract the year.
year = format(pr_bd, "%Y")
year = as.numeric(year)

iqr = IQR(prison_data$weight, na.rm = TRUE)
#summary(prison_data$weight)
#210+1.5*iqr
#165-1.5*iqr

iqr = IQR(prison_data$height, na.rm = TRUE)
#summary(prison_data$height, na.rm = TRUE)
#61-1.5*iqr
#72+1.5*iqr

prison <- prison_data %>% 
  filter(year >=1995 & year <1997) %>%
  filter(weight > 97.5) %>%
  filter(height > 53.5) 

#unique(prison_data$offender_status)
par_abs <- prison_data %>%
  filter(prison_data$offender_status == "PAROLE" | prison_data$offender_status == "ABSCONDER")


at_bd <- as.Date(athletics_data$dob, format="%Y-%m-%d")
year = format(at_bd, "%Y")
year = as.numeric(year)

athletics <- athletics_data %>%
  filter(year < 2003)
```

## r.tests
```{r r.tests}
weight_av = aov(lm(prison$BMI~prison$Zodaic_sign))
summary(weight_av)
#TukeyHSD(weight_av)

pr_table = table(droplevels(par_abs$offender_status),droplevels(par_abs$Zodaic_sign))
chisq.test(pr_table, correct = FALSE)

gold_zod<-aggregate(athletics$gold, by=list(Category=athletics$Zodaic_sign), FUN=sum)
names(gold_zod)[2] <- "Gold"
silver_zod<-aggregate(athletics$silver, by=list(Category=athletics$Zodaic_sign), FUN=sum)
names(silver_zod)[2] <- "Silver"
bronze<- aggregate(athletics$bronze, by=list(Category=athletics$Zodaic_sign), FUN=sum)
names(bronze)[2] <- "bronze"

gold_silv <- merge(gold_zod,silver_zod, by=c("Category"))
Trophies <- merge(gold_silv, bronze, by = c("Category"))
names(Trophies)[1] <- 'Zodiac Signs'
chisq.test(as.matrix(Trophies[,-1]), correct = FALSE)
```

## r tables and graphs
```{r tables and graphs}

ggplot(prison, aes(x=prison$Zodaic_sign, y=prison$BMI, fill=prison$Zodaic_sign))+
  geom_boxplot()+ coord_flip() + ggtitle("Zodiac signs Vs BMI") +
  xlab("Zodiac signs") + ylab("Body Mass Index(lb/inches^2)")+labs(fill = "Zodiac signs")

for_matrix = athletics %>% select(Zodaic_sign, gold, silver, bronze)
df = as.data.frame(for_matrix)
mt <- as.matrix(Trophies[,-1])
rownames(mt) <- c("Aquaries", "Aries", "Cancer", "Capriconr", "Germini", "Leo", "Libra", "Pisces", "Saggittarius","Scorpio", "Taurus", "Virgo")
ggballoonplot(mt,  fill = "value", main= "Zodiac signs and Trophies")+
  gradient_fill(c("black", "yellow", color = "#0073C2FF"))

pr_table
```

# Conclusion-1
According to the boxplot, it looks like there is variation among the range of Body mass index of the zodiac signs. However, in order to reject the null hypothesis and accept the alternate hypothesis, we need to get a p-value of less than 0.05. From the anova test and Tukey HSD test, we got a p-values of 0.55 (>0.05) which indicate that there is no statistically significant differece in mean BMI of the Zodiac signs. Hence, we fail to reject the null hypohtesis.


# Conclusion-2
From the chi squared test we got a p-value of 0.37(>0.05). Hence, we fail to reject the null hypothesis which states that the prisoners status is independent of their zodiac signs.

# Conclussion -3
The chi squared test on Zodiac signs vs type of Trophies table showed a p-value of 0.7338 (again >0.05). Therefore, eventhough it looks like there is some variation among the number of medals won by people born in different times, it is not statistically significant for us to come to conclussion. Again, we fail to reject the null hypothesis.

In general, we do not have evidence to prove that there is any connection between Zodiac signs and physical appearance, behaviour and success (determination to work).
